---
editor_options:
  chunk_output_type: console
---

NOTES:

-   This file is *not* included in the [source package](https://r-pkgs.org/package-structure-state.html#source-package) because of the [`.nopurl` suffix in the
    filename](https://rpkg.dev/pkgpurl/reference/purl_rmd.html#-rmd-files-excluded-from-purling).

-   The chunks below have to be manually executed in order to regenerate the package data.

# Internal data

NOTES:

-   Although the datasets below are saved as ["internal data" in `R/sysdata.rda`](https://r-pkgs.org/data.html#data-sysdata), some of them might still be
    exported and documented (in the main `.Rmd` document), something [not explicitly
    mentioned](https://coolbutuseless.github.io/2018/12/10/r-packages-internal-and-external-data/) in the book [R
    Packages](https://r-pkgs.org/data.html#data-data).

## `pkg_mgr_software`

Define what R-related software is available via Homebrew, Chocolatey and Scoop.

```{r}
pkg_mgr_software <- list(
  "Git" = list(brew = list(url = "https://formulae.brew.sh/formula/git",
                           cmd = "brew install git"),
               scoop = list(url = "https://scoop.sh/#/apps?q=git",
                            cmd = "scoop install git"),
               choco = list(url = "https://community.chocolatey.org/packages/git",
                            cmd = "choco install git")),
  "Quarto" = list(brew = list(url = "https://formulae.brew.sh/cask/quarto",
                              cmd = "brew install --cask quarto")),
  "R" = list(brew = list(url = "https://formulae.brew.sh/formula/r",
                         cmd = "brew install r"),
             scoop = list(url = "https://scoop.sh/#/apps?q=R+statistical+computing",
                          cmd = "scoop install r"),
             choco = list(url = "https://community.chocolatey.org/packages/R",
                          cmd = "choco install r")),
  "RStudio" = list(brew = list(url = "https://formulae.brew.sh/cask/rstudio",
                               cmd = "brew install --cask rstudio"),
                   scoop = list(url = "https://scoop.sh/#/apps?q=rstudio",
                                cmd = "scoop bucket add extras; scoop install rstudio"),
                   choco = list(url = "https://community.chocolatey.org/packages/R.Studio",
                                cmd = "choco install r.studio"))
)
```

## Write data

Save all the small data objects as a single internal file `R/sysdata.rda`. Note that when documenting them, they must be explicitly `@export`ed to be available
to package users.

```{r}
usethis::use_data(pkg_mgr_software,
                  internal = TRUE,
                  overwrite = TRUE,
                  compress = "xz",
                  version = 3L)
```

# Exported data

NOTES:

-   When used internally, other than internal data, exported data must always be referred to by its explicit namespace, i.e. e.g.
    `salim::regex_spelling_normalization`.

## `regex_spelling_normalization`

Regular expression rules for spelling normalization tailored to projects of the [Centre for Democracy Studies Aarau (ZDA)](https://www.zdaarau.ch/).

```{r}
regex_spelling_normalization <-
  tibble::tibble(id = character(),
                 category = character(),
                 purpose = character(),
                 pattern = list(),
                 replacement = character()) %>%
  # harmonize spelling
  tibble::add_row(id = "title_fa",
                  category = "harmonize_spelling",
                  purpose = "harmonize spelling of the [_FOKUS Aargau_](https://fokus.ag/) project title",
                  pattern = list("Fokus(?=[ -]Aargau)"),
                  replacement = "FOKUS") %>%
  tibble::add_row(id = "title_smartvote",
                  category = "harmonize_spelling",
                  purpose = "harmonize spelling of the [_smartvote_](https://smartvote.ch/) VAA",
                  pattern = list("\\bSmartvote\\b"),
                  replacement = "smartvote")
```

## Write data

Save all the bigger data objects to separate files under `data/*.rda`. Note that when documenting them, they mustn't be explicitly `@export`ed since they're
already automatically exported and thus available to package users.

```{r}
usethis::use_data(regex_spelling_normalization,
                  internal = FALSE,
                  overwrite = TRUE,
                  compress = "xz",
                  version = 3L)
```
