---
editor_options:
  chunk_output_type: console
---

# NOTES

-   This file is *not* included in the [source package](https://r-pkgs.org/structure.html#sec-source-package) because of the [`.nopurl` suffix in its
    filename](https://pkgpurl.rpkg.dev/reference/purl_rmd.html#-rmd-files-excluded-from-purling).

-   The chunks below have to be manually executed in order to regenerate the package data.

-   When used internally, other than internal data, [exported data](https://r-pkgs.org/data.html#sec-data-data) must always be referred to by its explicit namespace.

# Define data

## `regex_spelling_normalization`

Regular expression rules for spelling normalization tailored to projects of the [Centre for Democracy Studies Aarau (ZDA)](https://www.zdaarau.ch/).

```{r}
regex_spelling_normalization <-
  tibble::tibble(id = character(),
                 category = character(),
                 purpose = character(),
                 pattern = list(),
                 replacement = character()) %>%
  # harmonize spelling
  tibble::add_row(id = "title_fa",
                  category = "harmonize_spelling",
                  purpose = "harmonize spelling of the [_FOKUS Aargau_](https://fokus.ag/) project title",
                  pattern = list("Fokus(?=[ -]Aargau)"),
                  replacement = "FOKUS") %>%
  tibble::add_row(id = "title_smartvote",
                  category = "harmonize_spelling",
                  purpose = "harmonize spelling of the [_smartvote_](https://smartvote.ch/) VAA",
                  pattern = list("\\bSmartvote\\b"),
                  replacement = "smartvote")
```

# Write data

Save all the bigger data objects to separate files under `data/*.rda`. Note that when documenting them, they mustn't be explicitly `@export`ed since they're
already automatically exported and thus available to package users.

```{r}
usethis::use_data(regex_spelling_normalization,
                  internal = FALSE,
                  overwrite = TRUE,
                  compress = "xz",
                  version = 3L)
```
