---
editor_options:
  chunk_output_type: console
---

# INTERNAL

## Avoid `R CMD check` NOTES about undefined global objects

cf. <https://github.com/tidyverse/magrittr/issues/29#issuecomment-74313262>.

```{r}
utils::globalVariables(names = c(".",
                                 ":=",
                                 # other
                                 "download_url",
                                 "filename",
                                 "id",
                                 "name",
                                 "version_nr"))
```

## Functions

### `pkg_mgr_hint`

```{r}
pkg_mgr_hint <- function(software = names(pkg_mgr_software)) {
  
  software <- rlang::arg_match(software)
  
  result <- ""
  
  if (xfun::is_macos()) {
    result %<>% paste0(pkg_mgr_prose(software = software,
                                     pkg_mgrs = "brew"))
    
  } else if (xfun::is_windows()) {
    result %<>% paste0(pkg_mgr_prose(software = software,
                                     pkg_mgrs = c("scoop", "choco")))
  }
  
  result
}
```

### `pkg_mgr_prose`

```{r}
pkg_mgr_prose <- function(software,
                          pkg_mgrs = c("brew", "scoop", "choco")) {
  pkg_mgr_names <-
    pkg_mgrs |>
    dplyr::case_match("brew" ~ "Homebrew",
                      "scoop" ~ "Scoop",
                      "choco" ~ "Chocolatey")
  
  pkg_mgr_software |>
    purrr::chuck(software) |>
    purrr::keep_at(at = pkg_mgrs) |>
    purrr::map2_chr(.y = pkg_mgr_names,
                    .f = \(x, y) paste0(y, " ({.code ", x$cmd, "})")) |>
    pal::enum_str(sep2 = " or ") |>
    pal::when(length(.) > 0L ~ paste0(" ", software, " is also available via ", ., "."),
              ~ "")
}
```

# EXPORTED

## Spoken language

### Multi-language

#### `prettify_date`

DEPRECATED: `stringi::stri_datetime_format(format = c("date_short", "date_medium", "date_long", "date_full"), locale = "de-CH")` is superior in every regard!

TODO:

-   This fn somehow acts very strangely during FOKUS questionnaire generation: Although we use `salim::prettify_date(ballot_date, locale = 'de')`, it results in
    `October 18, 2020`, i.e. doesn't even use the proper format string.

NOTES:

-   Relevant SO question: [What is a reliable way of getting allowed locale names in
    R?](https://stackoverflow.com/questions/20960821/what-is-a-reliable-way-of-getting-allowed-locale-names-in-r)

-   Base R's conversion specifiers like `%Y` are documented in `?strptime`.

```{r}
#' Prettify date
#'
#' @description
#' `r lifecycle::badge("deprecated")`
#' 
#' Use [stringi::stri_datetime_format()] instead:
#' 
#' ```r
#' stringi::stri_datetime_format(format = c("date_short", "date_medium", "date_long", "date_full"),
#'                               locale = "en")
#' ```
#'
#' @details
#' Note that `prettify_date()` might only work on (Ubuntu) Linux in the current form since locales are one bitchy hell of a PITA...
#'
#' @param date Date to be prettified. A [date][base::Date] or something coercible to.
#' @param locale Locale the date should be prettified for. Currently only `"en"`/`"en-US"` and `"de"`/`"de-CH"` are implemented.
#'
#' @return A character scalar.
#' @family spoken
#' @keywords internal
#'
#' @examples
#' salim:::prettify_date("2021-12-21")
prettify_date <- function(date,
                          locale = c("en", "de", "en-US", "de-CH")) {

  lifecycle::deprecate_warn(when = "0.0.9015",
                            what = "prettify_date()",
                            details = paste0('Use the more powerful and more robust `stringi::stri_datetime_format(format = c("date_short", "date_medium", ',
                                             '"date_long", "date_full"), locale = "en")` function instead.'))
  
  locale <- rlang::arg_match(locale)

  withr::with_locale(new = c(LC_TIME = pal::when(. = locale,
                                                 . %in% c("en", "en-US") ~ "C",
                                                 . %in% c("de", "de-CH") ~ "de_CH.utf8")),
                     code =
                       locale |>
                       pal::when(. %in% c("en", "en-US") ~
                                   "%B %d, %Y",
                                 . %in% c("de", "de-CH") ~
                                   "%d. %B %Y",
                                 ~ cli::cli_abort("Specified {.arg locale} not implemented yet.")) |>
                       format(x = lubridate::as_date(date)))
}
```

#### `prettify_datetime`

DEPRECATED, see above.

```{r}
#' Prettify datetime
#'
#' @description
#' `r lifecycle::badge("deprecated")`
#' 
#' Use [stringi::stri_datetime_format()] instead:
#' 
#' ```r
#' stringi::stri_datetime_format(format = c("datetime_short", "datetime_medium", "datetime_long", "datetime_full"),
#'                               locale = "en")
#' ```
#'
#' @details
#' Note that `prettify_datetime()` might only work on (Ubuntu) Linux in the current form since locales are one bitchy hell of a PITA...
#'
#' @param datetime Datetime to be prettified. A [datetime][base::DateTimeClasses] or something coercible to.
#' @param locale Locale the datetime should be prettified for. Currently only `"en"`/`"en-US"` and `"de"`/`"de-CH"` are implemented.
#'
#' @return A character scalar.
#' @family spoken
#' @keywords internal
#'
#' @examples
#' salim:::prettify_datetime("2021-12-21T00:00:01Z")
prettify_datetime <- function(datetime,
                              locale = c("en", "de", "en-US", "de-CH")) {

  lifecycle::deprecate_warn(when = "0.0.9015",
                            what = "prettify_datetime()",
                            details = paste0('Use the more powerful and more robust `stringi::stri_datetime_format(format = c("datetime_short", ',
                                             '"datetime_medium", "datetime_long", "datetime_full"), locale = "en")` function instead.'))

  locale <- rlang::arg_match(locale)

  withr::with_locale(new = c(LC_TIME = pal::when(. = locale,
                                                 . %in% c("en", "en-US") ~ "C",
                                                 . %in% c("de", "de-CH") ~ "de_CH.utf8")),
                     code =
                       locale |>
                       pal::when(. %in% c("en", "en-US") ~
                                   "%B %d, %Y, %I:%M %p",
                                 . %in% c("de", "de-CH") ~
                                   "%d. %B %Y, %H:%M Uhr",
                                 ~ cli::cli_abort("Specified {.arg locale} not implemented yet.")) |>
                       format(x = lubridate::as_datetime(datetime)))
}
```

#### `rank_nr`

```{r}
#' Convert an integer into spelled abbreviated English or German rank
#'
#' @param x An integer scalar or something coercible to.
#' @param lang The language to write the rank in.
#'
#' @return A character scalar.
#' @family spoken
#' @export
rank_nr <- function(x,
                    lang = c("en", "de")) {
  
  lang <- rlang::arg_match(lang)
  result <- "?"
  
  if (length(x) > 0L) {
    
    if (lang == "de") {
      result <- switch(EXPR = as.character(as.integer(x)),
                       "1" = "erste",
                       "2" = "zweite",
                       "3" = "dritte",
                       "4" = "vierte",
                       "5" = "f\u00fcnfte",
                       "6" = "sechste",
                       "7" = "siebte",
                       "8" = "achte",
                       "9" = "neunte",
                       "?")
      
    } else if (lang == "en") {
      result <- switch(EXPR = as.character(as.integer(x)),
                       "1" = "1st",
                       "2" = "2nd",
                       "3" = "3rd",
                       "4" = "4th",
                       "5" = "5th",
                       "6" = "6th",
                       "7" = "7th",
                       "8" = "8th",
                       "9" = "9th",
                       "?")
    }
  }
  
  result
}
```

#### `write_out_n`

Caveats:

-   German is only covered from 0 to 9.

-   No support for grammatical gender -- articles of definite descriptors are always feminine.

```{r}
#' Write out a count
#'
#' @param n The count to write out in letters. A non-negative integer scalar.
#' @param lang The language to write out `n` in.
#' @param use_singular Use a definite descriptor for `n = 0` and `n = 1` instead of the indistinct number. A logical scalar.
#' @param use_dual Use a definite descriptor for `n = 2` instead of the indistinct number. A logical scalar.
#' @param strip_article Strip the definite article from the definite descriptors. Only relevant if at least one of `use_singular` or `use_dual` is `TRUE`. A
#'   logical scalar.
#'
#' @return A character scalar.
#' @seealso [gt::vec_fmt_spelled_num()] which spells out integers from 0 to 100 in `r length(setdiff(colnames(gt:::spelled_num), "number"))` different locales.
#' @family spoken
#' @export
write_out_n <- function(n,
                        lang = c("en", "de"),
                        use_singular = FALSE,
                        use_dual = FALSE,
                        strip_article = use_dual) {
  
  n %<>% checkmate::assert_count() %>% as.character()
  lang <- rlang::arg_match(lang)
  checkmate::assert_flag(use_singular)
  checkmate::assert_flag(use_dual)
  checkmate::assert_flag(strip_article)
  
  if (lang == "en") {
    
    result <- switch(EXPR = n,
                     `0` = ifelse(use_singular,
                                  "no",
                                  "zero"),
                     `1` = ifelse(use_singular,
                                  "the",
                                  "one"),
                     `2` = ifelse(use_dual,
                                  "the two",
                                  "two"),
                     xfun::numbers_to_words(as.integer(n)))
    
  } else if (lang == "de") {
    
    result <- switch(EXPR = n,
                     `0` = ifelse(use_singular,
                                  "keine",
                                  "null"),
                     `1` = ifelse(use_singular,
                                  "die",
                                  "ein"),
                     `2` = ifelse(use_dual,
                                  "die beiden",
                                  "zwei"),
                     `3` = "drei",
                     `4` = "vier",
                     `5` = "f\u00fcnf",
                     `6` = "sechs",
                     `7` = "sieben",
                     `8` = "acht",
                     `9` = "neun")
  }
  
  if (strip_article) {
    result %<>% stringr::str_remove(pattern = switch(EXPR = lang,
                                                     de = "^[Dd]ie ?",
                                                     en = "^[Tt]he ?"))
  }
  
  result
}
```

### English

#### `n_ify`

```{r}
#' Append `"n"` to indefinite article where indicated
#'
#' Appends an `"n"` to the indefinite article "a" if required by English language rules.
#'
#' Note that this function only covers part of all cases where "an" instead of "a" is indicated. Since the choice of _a_ or _an_ [depends on sound, not 
#' spelling](http://www.butte.edu/departments/cas/tipsheets/grammar/articles.html), it's not trivial to implement a universal solution.
#'
#' For example, this function doesn't properly append an `"n"` to the indefinite article in the following situations:
#' - A word starting with the letter "u" where it is pronounced as a vowel, e.g. _**unhappy**_. Note that most English words starting with letter "u" are
#'   pronounced differently like _universum_ and thus require the indefinite article _a_, not _an_.
#' - A word starting with a consonant that is pronounced as a vowel, e.g. _**hour**_.
#'
#' @param x A character vector.
#'
#' @return A character vector.
#' @family spoken
#' @export
#'
#' @examples
#' n_ify("This is a absolute disaster, such a error.")
n_ify <- function(x) {
  
  purrr::map_chr(.x = x,
                 .f = \(s) stringr::str_replace_all(string = s,
                                                    pattern = "(^[Aa]| a)(?= [AaEeOoIi]\\w+)",
                                                    replacement = "\\1n"))
}
```

### German

#### `definite_article_de`

```{r}
#' Get the German definite article
#'
#' @param gender The grammatical gender of the subject the definite article refers to.
#' @param case The grammatical case of the subject the definite article refers to.
#' @param pluralize Whether to return the plural form of the definite article. A logical scalar.
#'
#' @return A character scalar.
#' @family spoken
#' @export
definite_article_de <- function(gender = c("feminine",
                                           "masculine",
                                           "neuter"),
                                case = c("nominative",
                                         "genitive",
                                         "dative",
                                         "accusative"),
                                pluralize = FALSE) {
  
  gender <- rlang::arg_match(gender)
  case <- rlang::arg_match(case)
  
  if (checkmate::assert_flag(pluralize)) {
    
    switch(EXPR = case,
           nominative = "die",
           genitive = "der",
           dative = "den",
           accusative = "die")
    
  } else {
    
    switch(EXPR = case,
           nominative = switch(EXPR = gender,
                               feminine = "die",
                               masculine = "der",
                               neuter = "das"),
           genitive = switch(EXPR = gender,
                             feminine = "der",
                             "des"),
           dative = switch(EXPR = gender,
                           feminine = "der",
                           "dem"),
           accusative = switch(EXPR = gender,
                               feminine = "die",
                               masculine = "den",
                               neuter = "das"))
  }
}
```

#### `definite_article_de_declined`

```{r}
#' Decline the German definite article
#'
#' The declination of the German definite article [is a relative pronoun](https://de.wikipedia.org/wiki/Relativpronomen#Der,_die,_das).
#'
#' Note that the second singular feminine and plural genitive form `"derer"` is ignored by this function.
#'
#' @inheritParams definite_article_de
#'
#' @inherit definite_article_de return
#' @family spoken
#' @export
definite_article_de_declined <- function(gender = c("feminine",
                                                    "masculine",
                                                    "neuter"),
                                         case = c("nominative",
                                                  "genitive",
                                                  "dative",
                                                  "accusative"),
                                         pluralize = FALSE) {
  gender <- rlang::arg_match(gender)
  case <- rlang::arg_match(case)
  
  if (checkmate::assert_flag(pluralize)) {
    
    switch(EXPR = case,
           nominative = "die",
           genitive = "deren",
           dative = "deren",
           accusative = "die")
    
  } else {
    
    switch(EXPR = case,
           nominative = switch(EXPR = gender,
                               feminine = "die",
                               masculine = "der",
                               neuter = "das"),
           genitive = switch(EXPR = gender,
                             feminine = "deren",
                             "dessen"),
           dative = switch(EXPR = gender,
                           feminine = "der",
                           "dem"),
           accusative = switch(EXPR = gender,
                               feminine = "die",
                               masculine = "den",
                               neuter = "das"))
  }
}
```

#### `add_definite_article_de`

```{r}
#' Add the German definite article in dative case to a preposition
#'
#' @param preposition The preposition to add the German definite article in dative case to. A character scalar.
#' @param gender The grammatical gender of the `preposition`.
#'
#' @inherit definite_article_de return
#' @family spoken
#' @export
add_definite_article_de <- function(preposition,
                                    gender = c("feminine",
                                               "masculine",
                                               "neuter")) {
  checkmate::assert_string(preposition)
  gender <- rlang::arg_match(gender)
  
  if (gender == "feminine") {
    
    preposition |> pal::when(stringr::str_to_lower(.) == "zu" ~ paste0(., "r"),
                             ~ paste0(., " ", definite_article_de(gender = "feminine",
                                                                  case = "dative")))
  } else {
    
    preposition |> pal::when(stringr::str_to_lower(.) %in% c("bei", "zu") ~ paste0(., "m"),
                             stringr::str_to_lower(.) %in% c("in", "an") ~ stringr::str_replace(string = .,
                                                                                                pattern = "n$",
                                                                                                replacement = "m"),
                             ~ paste0(., " ", definite_article_de(gender = "masculine",
                                                                  case = "dative")))
  }
}
```

#### `decline_noun_de`

NOTES:

-   We expect `noun` to be declined in "indefinite" *nominative*, i.e. e.g. "weisser Mann", not "weisse Mann".
-   The function is very incomplete and just covers the cases we encountered during development of the fokus pkg.

```{r}
#' Decline a German noun
#'
#' Note that this function is far from covering all grammatical possibilities. I.a. it currently doesn't support the [special
#' cases](https://deutsch.lingolia.com/de/grammatik/nomen/deklination/genitiv#a-besonderheiten) where a masculine noun gets an `n` or `en` ending attached in
#' the accusative, dative and genitive cases.
#'
#' @param noun The noun to decline, optionally with preceding adjective(s), in (indefinite) nominative. A character scalar.
#' @param gender The grammatical gender of the `noun`.
#' @param case The grammatical case of the `noun`.
#' @param is_plural Whether or not `noun` is a plural.
#'
#' @inherit definite_article_de return
#' @family spoken
#' @export
decline_noun_de <- function(noun,
                            gender = c("feminine",
                                       "masculine",
                                       "neuter"),
                            case = c("nominative",
                                     "genitive",
                                     "dative",
                                     "accusative"),
                            is_plural = FALSE) {
  
  gender <- rlang::arg_match(gender)
  case <- rlang::arg_match(case)
  checkmate::assert_flag(is_plural)
  
  result <- noun
  
  if (gender != "feminine" && case == "genitive") {
    
    result %<>% pal::when(stringr::str_detect(string = ., pattern = "(ss|\U00DF|x|z)$") ~ paste0(., "es"),
                          endsWith(., "s") ~ paste0(., "ses"),
                          ~ paste0(., "s"))
    
    # decline preceeding adjectives
  } else if (case %in% c("genitive", "dative") || (case == "accusative" && is_plural)) {
    result %<>% stringr::str_replace_all(pattern = "((?:^|[:blank:])[:alnum:]+e)([:blank:])(?!\\s*$)",
                                         replacement = "\\1n\\2")
  }
  
  result
}
```

## Regular expression rules

### `regex_spelling_normalization`

```{r}
#' Regular expression patterns and replacements for spelling normalization
#'
#' @format `r pkgsnip::return_lbl("tibble")`
#' @seealso
#' Other regular expression rules: [`yay::regex_text_normalization`], [`yay::regex_file_normalization`]
#' 
#' String normalization functions: [yay::str_normalize()], [yay::str_normalize_file()]
#'
#' @examples
#' # unnest the pattern column
#' salim::regex_spelling_normalization |> tidyr::unnest_longer(col = pattern)
"regex_spelling_normalization"
```

## Pandoc

TODO: Move these fns to dedicated pkg *pndc*? Then the pandoc templates at [gitlab.com/salim_b/pandoc/templates](https://gitlab.com/salim_b/pandoc/templates)
should be moved inside the pndc pkg repo, but build-ignored.

### `download_pandoc_binaries`

```{r}
#' Download Pandoc release
#'
#' Downloads the [assets][pandoc_release_assets] and extracts the executable program binaries for Linux, macOS and Windows of the specified Pandoc release.
#'
#' @inheritParams pandoc_release_assets
#' @param os The operating system(s) for which Pandoc binaries should be downloaded. Any combination of
#'   `r pal::fn_param_defaults(param = "os", fn = download_pandoc_binaries) |> pal::wrap_chr("\x60") |> cli::ansi_collapse(last = " and ")`.
#' @param path The filesystem path to which the binaries are saved to. A [path][fs::fs_path] or something coercible to.
#' @param overwrite Whether to overwrite existing binaries under `path`.
#'
#' @return `path` invisibly.
#' @family pandoc
#' @seealso GitHub release functions: [yay::gh_releases()], [yay::gh_release_latest()]
#' @export
download_pandoc_binaries <- function(release_id = yay::gh_release_latest(owner = "jgm",
                                                                         name = "pandoc")$id,
                                     os = c("linux", "macos", "windows"),
                                     path = "bin/",
                                     overwrite = TRUE) {
  rlang::check_installed("yay",
                         reason = pal::reason_pkg_required())
  rlang::check_installed("waldo",
                         reason = pal::reason_pkg_required())
  checkmate::assert_subset(os,
                           choices = eval(formals()$os),
                           empty.ok = FALSE)
  checkmate::assert_path_for_output(path,
                                    overwrite = TRUE)
  checkmate::assert_flag(overwrite)
  
  path_tmp <- fs::path_temp()
  
  assets <-
    pandoc_release_assets(release_id = release_id) |>
    dplyr::filter(stringr::str_detect(string = filename,
                                      pattern = paste0("(?i)", pal::fuse_regex(os), "(-(amd64|x86_64))?\\.(zip|tar\\.gz)"))) |>
    dplyr::mutate(download_path = fs::path(path_tmp, filename))
  
  # download assets to tmp dir
  purrr::walk2(.x = assets$download_url,
               .y = assets$download_path,
               .f = \(url, path) utils::download.file(url = url,
                                                      destfile = path,
                                                      method = "auto",
                                                      mode = "wb",
                                                      cacheOK = TRUE,
                                                      quiet = TRUE))
  # extract binaries to tmp dir
  assets |>
    dplyr::select(-download_url) |>
    as.list() |>
    purrr::pwalk(.f = \(filename, os, download_path) {
      
      path_os_tmp <- fs::path(path_tmp, "pandoc", os)
      
      # extract all files flat
      if (endsWith(filename, ".zip")) {
        
        zip::unzip(zipfile = download_path,
                   exdir = path_os_tmp,
                   junkpaths = TRUE)
      } else {
        
        utils::untar(tarfile = download_path,
                     exdir = path_os_tmp)
        
        # since `untar()` has no "junkpaths"-like option, we flatten the dir structure ourselves
        fs::dir_ls(path = path_os_tmp,
                   type = "file",
                   recurse = TRUE,
                   all = TRUE) |>
          fs::file_move(new_path = path_os_tmp)
      }
      
      # delete archive and all extracted files but the Pandoc binary
      fs::file_delete(download_path)
      
      fs::dir_ls(path = path_os_tmp,
                 regexp = "[\\/]pandoc(\\.exe)?$",
                 all = TRUE,
                 invert = TRUE) |>
        fs::file_delete()
    })
  
  # move binaries to final dir
  fs::dir_create(path = fs::path(path, "pandoc"))
  
  path_tmp |>
    fs::path("pandoc") %T>%
    fs::dir_copy(new_path = fs::path(path, "pandoc"),
                 overwrite = overwrite) %>%
    fs::dir_delete()
  
  invisible(path)
}
```

### `pandoc_release_assets`

```{r}
#' List Pandoc release assets
#'
#' Uses [gh::gh()] to fetch filenames, corresponding operating systems and download URLs of a specific [GitHub
#' release](https://docs.github.com/repositories/releasing-projects-on-github) of [Pandoc](https://pandoc.org/) via [GitHub's REST
#' API](https://docs.github.com/en/rest/reference/repos#list-release-assets) and returns them as a [tibble][tibble::tbl_df].
#'
#' @param release_id The GitHub release identifier of the desired Pandoc release. Use
#'   [`yay::gh_releases(owner = "jgm", name = "pandoc")`][yay::gh_releases] to determine the release identifier of a specific Pandoc version number. An
#'   integer scalar.
#'
#' @return `r pkgsnip::param_lbl("tibble")`
#' @family pandoc
#' @seealso GitHub release functions: [yay::gh_releases()], [yay::gh_release_latest()]
#' @export
pandoc_release_assets <- function(release_id = yay::gh_release_latest(owner = "jgm",
                                                                      name = "pandoc")$id) {
  rlang::check_installed("yay",
                         reason = pal::reason_pkg_required())
  rlang::check_installed("gh",
                         reason = pal::reason_pkg_required())
  checkmate::assert_count(release_id,
                          positive = TRUE)
  
  gh::gh(endpoint = "/repos/{owner}/{repo}/releases/{release_id}/assets", # nolint
         owner = "jgm",
         repo = "pandoc",
         release_id = release_id,
         .method = "GET") |>
    purrr::map(\(x) tibble::tibble(filename = x$name,
                                   os =
                                     x$name |>
                                     stringr::str_extract(pattern = "(?i)(linux|macos|windows|\\.deb$)") |>
                                     stringr::str_to_lower() |>
                                     stringr::str_replace(pattern = stringr::fixed(".deb"),
                                                          replacement = "linux"),
                                   download_url = x$browser_download_url)) |>
    purrr::list_rbind()
}
```

### `pandoc_tpl`

```{r}
#' Download Pandoc template
#'
#' Downloads one of our [custom Pandoc templates](https://gitlab.com/salim_b/pandoc/templates) and returns it as a character scalar. The template is cached
#' using [pal::req_cached()], thus only re-downloaded if the local copy is outdated.
#'
#' @param tpl Template to download. One of
#'   `r pal::fn_param_defaults(param = "tpl", fn = "pandoc_tpl") |> pal::wrap_chr("\x60") |> cli::ansi_collapse(last = " and ")`.
#'
#' @return The template as a character scalar.
#' @family pandoc
#' @export
#' 
#' @examples
#' \dontrun{
#' salim::pandoc_tpl(tpl = "quarto_mod.latex") |>
#'   brio::write_file(path = "quarto_mod.latex")}
pandoc_tpl <- function(tpl = "quarto_mod.latex") {
  
  tpl <- rlang::arg_match(tpl)
  
  (pal::req_cached(url = glue::glue("https://gitlab.com/salim_b/pandoc/templates/-/raw/main/{tpl}?ref_type=heads&inline=false")) |>
      httr2::resp_body_string())
}
```

## Development environment currentness

### `lvl_up_r`

```{r}
#' Level up R
#'
#' Checks whether the installed R version is >= a cached reference version. Intended to level up the R version in use between multiple users of the same code
#' (e.g. contributors to a specific R project).
#'
#' If the file `path_min_vrsn` exists, it is checked whether the installed R version is greater than or equal to the version number stored in that file, and if
#' not, an alert is displayed.
#' 
#' If `update_min_vrsn` is set to `TRUE` and the file `path_min_vrsn` doesn't exist or contains an R version string that's lower than the currently installed
#' version of R, `path_min_vrsn` is overwritten with the currently installed R version string.
#'
#' @param path_min_vrsn Path to the cached R version string.
#' @param update_min_vrsn Whether or not to overwrite `path_min_vrsn` with the currently installed R version string *iff* the latter is higher than the former.
#'
#' @return Currently installed R version as a [numeric version][as.package_version], invisibly.
#' @family dev_env
#' @export
lvl_up_r <- function(path_min_vrsn,
                     update_min_vrsn = FALSE) {
  
  checkmate::assert_flag(update_min_vrsn)
  
  if (fs::file_exists(path_min_vrsn)) {
    
    min_vrsn <-
      path_min_vrsn |>
      brio::read_lines(n = 1L) |>
      as.package_version()
    
  } else {
    min_vrsn <- as.package_version("0.0")
  }
  
  current_vrsn <-
    R.Version() %$%
    glue::glue("{major}.{minor}") |>
    as.package_version()
  
  if (isTRUE(current_vrsn < min_vrsn)) {
    
    cli::cli_alert_warning(paste0(
      "Your version of {.pkg R} is out of date. Please update to version {.val {min_vrsn}} or above. The latest stable release is available at ",
      "{.url https://cloud.r-project.org/}.", pkg_mgr_hint("R")
    ))
  }
  
  if (update_min_vrsn && current_vrsn > min_vrsn) {
    
    brio::write_lines(text = current_vrsn,
                      path = path_min_vrsn)
  }
  
  invisible(current_vrsn)
}
```

### `lvl_up_rstudio`

```{r}
#' Level up RStudio
#'
#' Checks whether the currently running [RStudio](https://rstudio.com/products/rstudio/) version is >= a cached reference version. Intended to level up the 
#' RStudio version in use between multiple users of the same code (e.g. contributors to a specific R project).
#'
#' If the file `path_min_vrsn` exists, it is checked whether RStudio is running, and if so, whether its version number is greater than or equal to the version
#' number stored in that file. If this is not the case, an alert is displayed.
#' 
#' If `update_min_vrsn` is set to `TRUE` and the file `path_min_vrsn` doesn't exist or contains an RStudio version string that's lower than the currently
#' running version of RStudio, `path_min_vrsn` is overwritten with the currently running RStudio version string.
#'
#' Note that since **RStudio must be running for `lvl_up_rstudio()` to have any effect**, it can't be directly called in a (project-level)
#' [`.Rprofile`](https://support.rstudio.com/hc/en-us/articles/360047157094-Managing-R-with-Rprofile-Renviron-Rprofile-site-Renviron-site-rsession-conf-and-repos-conf)
#' startup script. Instead, the [`rstudio.sessionInit`](https://docs.rstudio.com/ide/server-pro/r_sessions/session_startup_scripts.html) hook can be used.
#' Example:
#' 
#' ```r
#' setHook(hookName = "rstudio.sessionInit",
#'         value = function(newSession) {
#'             if (newSession)
#'                 salim::lvl_up_rstudio(path_min_vrsn = ".RStudio_version",
#'                                       update_min_vrsn = TRUE)
#'         },
#'         action = "append")
#' ```
#'
#' @param path_min_vrsn Path to the cached RStudio version string.
#' @param update_min_vrsn Whether or not to overwrite `path_min_vrsn` with the currently running RStudio version string *iff* the latter is higher than the
#'   former.
#'
#' @return Currently running RStudio version as a [numeric version][as.package_version], or `NULL` if RStudio is not running, invisibly.
#' @family dev_env
#' @export
lvl_up_rstudio <- function(path_min_vrsn,
                           update_min_vrsn = FALSE) {
  
  checkmate::assert_flag(update_min_vrsn)
  rlang::check_installed("rstudioapi",
                         reason = pal::reason_pkg_required())
  current_vrsn <- NULL
  
  # skip if RStudio is not running
  if (rstudioapi::isAvailable()) {
    
    if (fs::file_exists(path_min_vrsn)) {
      
      min_vrsn <-
        path_min_vrsn |>
        brio::read_lines(n = 1L) |>
        as.package_version()
      
    } else {
      min_vrsn <- as.package_version("0.0")
    }
    
    current_vrsn <-
      rstudioapi::versionInfo() |>
      purrr::chuck("version")
    
    if (isTRUE(current_vrsn < min_vrsn)) {
      
      cli::cli_alert_warning(paste0(
        "Your version of {.pkg RStudio} is out of date. Please update to version {.val {min_vrsn}} or above. The latest version is available at ",
        "{.url https://posit.co/download/rstudio-desktop/}.", pkg_mgr_hint("RStudio")
      ))
    }
    
    if (update_min_vrsn && current_vrsn > min_vrsn) {
      
      brio::write_lines(text = current_vrsn,
                        path = path_min_vrsn)
    }
  }
  
  invisible(current_vrsn)
}
```

### `lvl_up_quarto`

```{r}
#' Level up Quarto
#'
#' Checks whether the installed [Quarto](https://quarto.org/) version is >= a cached reference version. Intended to level up the Quarto version in use between
#' multiple users of the same code (e.g. contributors to a specific Quarto project).
#'
#' @param path_min_vrsn Path to the cached Quarto version string.
#' @param update_min_vrsn Whether or not to overwrite `path_min_vrsn` with the currently installed Quarto version string *iff* the latter is higher than the
#'   former.
#'
#' @return Currently installed Quarto version as a [numeric version][numeric_version], invisibly.
#' @family dev_env
#' @export
lvl_up_quarto <- function(path_min_vrsn,
                          update_min_vrsn = FALSE) {
  
  checkmate::assert_flag(update_min_vrsn)
  rlang::check_installed("quarto",
                         reason = pal::reason_pkg_required())
  
  if (fs::file_exists(path_min_vrsn)) {
    
    min_vrsn <-
      path_min_vrsn |>
      brio::read_lines(n = 1L) |>
      as.package_version()
    
  } else {
    min_vrsn <- as.package_version("0.0")
  }
  
  current_vrsn <-
    quarto::quarto_path() |>
    system2(args = "--version",
            stdout = TRUE,
            stderr = TRUE) |>
    as.package_version()
  
  if (isTRUE(current_vrsn < min_vrsn)) {
    
    cli::cli_alert_warning(paste0(
      "Your version of {.pkg Quarto} is out of date. Please update to version {.val {min_vrsn}} or above. The latest stable release is available at ",
      "{.url https://quarto.org/docs/get-started/}.", pkg_mgr_hint("Quarto")
    ))
  }
  
  if (update_min_vrsn && current_vrsn > min_vrsn) {
    
    brio::write_lines(text = current_vrsn,
                      path = path_min_vrsn)
  }
  
  invisible(current_vrsn)
}
```

### `update_rpkgs`

```{r}
#' Update R packages from rpkg.dev
#'
#' Installs/updates \R packages from [rpkg.dev](https://gitlab.com/rpkg.dev) to the latest (development) version.
#'
#' @param pkgs R pkgs to be updated. A subset of:
#'   `r pal::fn_param_defaults(fn = "update_rpkgs", param = "pkgs") |> pal::wrap_chr("\x60") |> pal::as_md_list()`
#' @param from_cran Whether or not to install `pkgs` from a [CRAN](https://en.wikipedia.org/wiki/R_package#Comprehensive_R_Archive_Network_(CRAN)) repository
#'   if possible. The repository URLs set in the [`repos`][options] \R option are used by default, with fallback to `https://cloud.r-project.org/`. If `FALSE`,
#'   the latest *development* versions are always installed, regardless of whether `pkgs` are available on CRAN or not.
#'
#' @return `pkgs`, invisibly.
#' @family dev_env
#' @export
update_rpkgs <- function(pkgs = c("nocodb",
                                  "pal",
                                  "pkgdown.tpl",
                                  "pkgpins",
                                  "pkgpurl",
                                  "pkgsnip",
                                  "plotlee",
                                  "qstnr",
                                  "rstd",
                                  "swissmuni",
                                  "tocr",
                                  "yay"),
                         from_cran = FALSE) {
  
  rlang::check_installed("remotes",
                         reason = pal::reason_pkg_required())
  pkgs <- rlang::arg_match(pkgs,
                           multiple = TRUE)
  purrr::walk(pkgs,
              \(x) {
                
                if (pal::is_pkg_cran(x)) {
                  utils::install.packages(pkgs = x,
                                          repos = getOption("repos",
                                                            default = "https://cloud.r-project.org/"))
                } else {
                  remotes::install_gitlab(repo = paste0("rpkg.dev/", x), # nolint: paste_linter
                                          upgrade = FALSE)
                }
              })
  
  invisible(pkgs)
}
```

### `update_salims_pkgs`

```{r}
#' Update Salim B's R packages
#'
#' Installs/updates all of [Salim B's R packages](https://gitlab.com/salim_b/r/pkgs) to the latest (development) version.
#'
#' @inheritParams update_rpkgs
#' @param pkgs R pkgs to be updated. A subset of:
#'   `r pal::fn_param_defaults(fn = "update_salims_pkgs", param = "pkgs") |> pal::wrap_chr("\x60") |> pal::as_md_list()`
#'
#' @return `pkgs`, invisibly.
#' @family dev_env
#' @export
update_salims_pkgs <- function(pkgs = c("salim",
                                        "swissparty"),
                               from_cran = FALSE) {
  
  rlang::check_installed("remotes",
                         reason = pal::reason_pkg_required())
  pkgs <- rlang::arg_match(pkgs,
                           multiple = TRUE)
  purrr::walk(pkgs,
              \(x) {
                
                if (pal::is_pkg_cran(x)) {
                  utils::install.packages(pkgs = x,
                                          repos = getOption("repos",
                                                            default = "https://cloud.r-project.org/"))
                } else {
                  remotes::install_gitlab(repo = paste0("salim_b/r/pkgs/", x), # nolint: paste_linter
                                          upgrade = FALSE)
                }
              })
  
  invisible(pkgs)
}
```

### `update_zdaarau_pkgs`

```{r}
#' Update R packages from the Centre for Democracy Studies Aarau (ZDA)
#'
#' Installs/updates all of the [R packages from the Centre for Democracy Studies Aarau (ZDA)](https://gitlab.com/zdaarau/rpkgs) to the latest (development)
#' version.
#'
#' @inheritParams update_rpkgs
#' @param pkgs R pkgs to be updated. A subset of:
#'   `r pal::fn_param_defaults(fn = "update_zdaarau_pkgs", param = "pkgs") |> pal::wrap_chr("\x60") |> pal::as_md_list()`
#'
#' @return `pkgs`, invisibly.
#' @family dev_env
#' @export
update_zdaarau_pkgs <- function(pkgs = c("rdb.report",
                                         "fokus",
                                         "rdb",
                                         "swissevote"),
                                from_cran = FALSE) {
  
  rlang::check_installed("remotes",
                         reason = pal::reason_pkg_required())
  pkgs <- rlang::arg_match(pkgs,
                           multiple = TRUE)
  checkmate::assert_flag(from_cran)
  
  purrr::walk(pkgs,
              \(x) {
                
                if (from_cran && pal::is_pkg_cran(x)) {
                  utils::install.packages(pkgs = x,
                                          repos = getOption("repos",
                                                            default = "https://cloud.r-project.org/"))
                } else {
                  remotes::install_gitlab(repo = paste0("zdaarau/rpkgs/", x), # nolint: paste_linter
                                          upgrade = FALSE)
                }
              })
  
  invisible(pkgs)
}
```
