---
editor_options:
  chunk_output_type: console
---

# NOTES

-   This file is *not* included in the [source package](https://r-pkgs.org/structure.html#sec-source-package) because of the [`.nopurl` suffix in its
    filename](https://pkgpurl.rpkg.dev/reference/purl_rmd.html#-rmd-files-excluded-from-purling).

-   The chunks below have to be manually executed in order to regenerate the package data.

-   Although the datasets below are saved as ["internal data" in `R/sysdata.rda`](https://r-pkgs.org/data.html#sec-data-sysdata), they can still be exported and
    documented (by documenting the dataset's quoted name in the main `.Rmd` source file – which only works when the dataset is also `@export`ed), something [not
    explicitly mentioned](https://coolbutuseless.github.io/2018/12/10/r-packages-internal-and-external-data/) in the book [R
    Packages](https://r-pkgs.org/data.html#sec-data-data). To do so, you first need to manually add the `export()` directive in the NAMESPACE file since
    roxygen2 won't add it automatically.

# Define data

## `pkg_mgr_software`

Define what R-related software is available via Homebrew, Chocolatey and Scoop.

```{r}
pkg_mgr_software <- list(
  "Git" = list(brew = list(url = "https://formulae.brew.sh/formula/git",
                           cmd = "brew install git"),
               scoop = list(url = "https://scoop.sh/#/apps?q=git",
                            cmd = "scoop install git"),
               choco = list(url = "https://community.chocolatey.org/packages/git",
                            cmd = "choco install git")),
  "Quarto" = list(brew = list(url = "https://formulae.brew.sh/cask/quarto",
                              cmd = "brew install --cask quarto")),
  "R" = list(brew = list(url = "https://formulae.brew.sh/formula/r",
                         cmd = "brew install r"),
             scoop = list(url = "https://scoop.sh/#/apps?q=R+statistical+computing",
                          cmd = "scoop install r"),
             choco = list(url = "https://community.chocolatey.org/packages/R",
                          cmd = "choco install r")),
  "RStudio" = list(brew = list(url = "https://formulae.brew.sh/cask/rstudio",
                               cmd = "brew install --cask rstudio"),
                   scoop = list(url = "https://scoop.sh/#/apps?q=rstudio",
                                cmd = "scoop bucket add extras; scoop install rstudio"),
                   choco = list(url = "https://community.chocolatey.org/packages/R.Studio",
                                cmd = "choco install r.studio"))
)
```

# Write data

Save all the small data objects as a single internal file `R/sysdata.rda`. Note that when documenting them, they must be explicitly `@export`ed to be available
to package users.

```{r}
usethis::use_data(pkg_mgr_software,
                  internal = TRUE,
                  overwrite = TRUE,
                  compress = "xz",
                  version = 3L)
```
